---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

Attempting to model the S&P 500 annual returns (including dividends).

Lets read in the file:

```{r}
library(tidyverse)
df_sp500 = readxl::read_excel('SP500_annual_returns.xlsx')
df_sp500 = df_sp500 %>% 
  rename(return = '% Return', year = 'Year') %>%
  mutate(year = as.integer(year))
```

Now lets log transform it as returns are multiplicative, i.e., a return of -50% is not the same as +50% - -50% and +100% are equivalent, which we can model by log transforming.

```{r}
df_sp500 = df_sp500 %>% mutate(log1p_return = log1p(return/100.0))
```

And do some plots:
```{r}
ggplot(data = df_sp500, aes(log1p_return)) + geom_histogram(bins = 25)
ggplot(data = df_sp500, aes(sample = log1p_return)) + stat_qq() + stat_qq_line()
```

Eyeballing it looks very highly skewed.

```{r}
library(moments)
skewness(df_sp500 %>% pull(log1p_return))
```
Let's calculate the moments, and use the sn package to create samples that return the log1p returns.

```{r}
sp500_moments = list(
  mean = mean(df_sp500 %>% pull(log1p_return)),
  sd = sd(df_sp500 %>% pull(log1p_return)),
  skew = skewness(df_sp500 %>% pull(log1p_return))
)
```


```{r}
library(sn)

sp_params = cp2dp(c(sp500_moments$mean, sp500_moments$sd, sp500_moments$skew), "SN")

df_returns = tibble(log1p_return = rsn(n=1000, dp = sp_params))

df_returns = df_returns %>% mutate(return = expm1(log1p_return))

```


And what about assuming a lower return going forward:

```{r}
lower_params = cp2dp(c(0.05, 0.12, -0.5), "SN")

df_returns = df_returns %>% mutate(log1p_lower_return = rsn(n=1000, dp = lower_params))

df_returns = df_returns %>% mutate(lower_return = expm1(log1p_lower_return))

df_returns
```

```{r}
ggplot(data = df_returns, aes(return)) + geom_histogram(bins = 100)

ggplot(data = df_returns, aes(lower_return)) + geom_histogram(bins = 100)
```
